---
output: 
  github_document:
    html_preview: FALSE
    keep_html: FALSE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# spotoroo

<!-- badges: start -->
[![R-CMD-check](https://github.com/TengMCing/spotoroo/workflows/R-CMD-check/badge.svg)](https://github.com/TengMCing/spotoroo/actions)
[![Codecov test coverage](https://codecov.io/gh/TengMCing/spotoroo/branch/master/graph/badge.svg)](https://app.codecov.io/gh/TengMCing/spotoroo?branch=master)
<!-- badges: end -->

## Overview 

"spotoroo" stands for spatiotemporal clustering in R of hot spot data. It is an algorithm to cluster satellite hot spots, detect ignition points and reconstruct fire movement.

## Installation

You can install the released version of spotoroo from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("spotoroo")
```

You can install the development version from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("TengMCing/spotoroo")
```
## Usage

```{r}
library(spotoroo)
```

The below examples use the built-in dataset `hotspots`. The hot spot data needs to has at least three columns: the longitude, the latitude, and the observed time.

```{r}
str(hotspots)
```

Perform spatiotemporal clustering on this dataset. You need to specify which columns correspond to the spatial variables ("lon", "lat"), and which to observed time ("obsTime"). 

There is a choice of options for the algorithm. 

- "activeTime" sets the time to consider that a fire can be active, and longer than this between hot spots will create a new cluster 

- "adjDist" sets the maximum intra-cluster spatial distance between nearest hot spots beyond which they would be considered part of a different cluster 

- "minPts" sets the minimum number of hot spots in a cluster

- "minTime" sets the minimum length of time of a cluster

- "ignitionCenter" sets the method to calculate the ignition points

- "timeUnit" and "timeStep" set the length of time between successive time indexes

```{r example}
result <- hotspot_cluster(hotspots,
                          lon = "lon",
                          lat = "lat",
                          obsTime = "obsTime",
                          activeTime = 24,
                          adjDist = 3000,
                          minPts = 4,
                          minTime = 3,
                          ignitionCenter = "mean",
                          timeUnit = "h",
                          timeStep = 1)


result
```


You can make a summary of the clustering results.

```{r}
summary(result)
```

You can extract a subset of clusters from the results.

```{r}
fire_1_and_2 <- extract_fire(result, 1:2)
head(fire_1_and_2, 2)
```


Plot of the result. In this example, there is a total of 6 clusters, so all can be displayed. 

```{r out.width = "70%", out.height="70%"}
plot(result, bg = plot_vic_map())
```

You can also choose a subset of clusters, and this will plot without a map, so that you can see a zoomed in view of the hot spot clusters and their ignition points. 

```{r out.width = "70%", out.height="70%"}
plot(result, cluster = c(1,2,3,4))
```

To examine the fire movements, use the option "mov", and the movement will be shown as connected lines between centroids at each time step, for each cluster.  

```{r out.width = "70%", out.height="70%"}
plot(result,
   type = "mov",
   cluster = 1:3,
   step = 6,
   bg = plot_vic_map())
```


To examine the time line of clusters and learn about the intensity of fire periods, use the option "timeline".

```{r out.width = "70%", out.height="70%"}
plot(result, "timeline", 
     dateLabel = "%b %d", 
     mainBreak = "1 week")
```

