---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hotspotcluster

<!-- badges: start -->
<!-- badges: end -->

The goal of hotspotcluster is to ...

## Installation

<!-- You can install the released version of hotspotcluster from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("hotspotcluster") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("TengMCing/hotspotcluster")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r echo = FALSE, message=FALSE, warning=FALSE, error=FALSE}
library(devtools)
load_all()
```

The first 20 observations of the built-in dataset `hotspots`.

```{r}
hotspotcluster::hotspots[1:20,]
```

Perform spatiotemporal clustering on this data.

```{r example}
results <- hotspotcluster::hotspot_cluster(hotspots)
```

The ignition points of the first 20 bushfires.

```{r}
data.frame(lon = results$ignition_lon, lat = results$ignition_lat)[1:20,]
```

The memberships of the first 20 hotspots.

```{r}
results$memberships[1:20]
```


